Developing a preprint culture in biology

### 2012-10-05 ms.tex
merge DGs file

\author{Philippe Desjardins-Proulx}
\email[E-mail: ]{philippe.d.proulx@gmail.com}
\affiliation{Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Canada.}
\affiliation{Quebec Center for Biodiversity Science, McGill University, Canada.}
\affiliation{D\'epartement des sciences biologiques, Universit\'e du Qu\'ebec \`a Montr\'eal, Canada.}

@@

Furthermore, the review process as a whole is critically over-loaded, because
as the number of active scientists increases, because the pressure to publish
increases, and because of an effect dubbed ``the tragedy of the reviewers
commons'' \cite{hoc09}.  At the same time, rejection rates are high in most journals
\cite{aar08,roh09}, and when not invited to submit a revision, authors must start the whole
process all over again.  Initiatives to reduce time from submission to
publication have emerged across the scientific community. Rohr et al.
\cite{roh09} called for the recycling and reuse of peer-reviews:
by attaching previous reviews and detailed replies to a new submission, both
the editor and the referees can gauge the work done on the manuscript, and
perhaps evaluate it with less prejudice. In a similar way, the \emph{Peerage
of Science} initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can
achieve the same goal of reducing the time spent in review. By putting a
manuscript for open comments and criticisms, the authors will receive valuable
feedback and can improve the version which will be submitted. With a rich
enough community of scientists depositing preprints, and commenting on them,
the process of an open pre- review can become widespread and will overall
increase the quality of first submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable.  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map} \end{figure}

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal
\cite{gin11}. Ironically, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.

Prepublication reviews by a small network of colleagues is an important part of
the scientific process, attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the scientific process can be open
as soon as possible.

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, Figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually
be reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
response. F1000Research works closely with data providers to integrate raw
data to the paper. Forinstance, upon submitting a paper, authors are asked to
upload their data, which are then integrated in \emph{e.g.} FigShare widgets,
the DOI of which are given in the paper when the data are first mentionned.
The licensing of the data is similar to the one used by FigShare, meaning that
the articles are free to access, and can be redistributed readily. By putting
much effort in integrating data to the paper, F1000Research is working to make
science more reproducible and open.

\subsection{Github}

This manuscript was developed entirely as an open project on github. github is
one of several hosting services for collaborative development using the git
version control system (VCS).  git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous contributers to work asynchronously on the same project, often in parallel branches, all of which can be effortlessly merged and version controlled. Although created primarily for software development, git is ideal for academic research since it provides a way to version control and collaborate on every step of the manuscript development process, from data manipulation and analysis to writing and revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then merge their changes into a master branch.This takes the preprint process to an entirely new level, where the entire writing process is open from the beginning.
% I would avoid using the term forking since it is not a central piece of the section. It is one specific way of collaborating on a github repo when everyone does not have rights to it. People could just as easily work on a single branch, all with write access, either directly to the master branch or to individual branches that could be merged. So in other words, we should keep the language general so people don't get bogged down with trying to understand what a fork is.
% also I changed one author clones, then submits pull request langauge a bit since the main point of using git is that we can all work simultaneously unlike svn.
% I took out "These techniques are widely used in open source software development." because git can also be used for closed source software, offering the same advantages, where the repositories are private. Paid git accounts allow for private repos.

@@

>>>>

\author{Philippe Desjardins-Proulx}
\email[E-mail: ]{philippe.d.proulx@gmail.com}
\affiliation{Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Canada.}
\affiliation{Quebec Center for Biodiversity Science, McGill University, Canada.}

@@

Furthermore, the review process as a whole is critically over-loaded, because
the number of active scientists increases, because the pressure to publish
increases, and because of an effect dubbed ``the tragedy of the reviewers
commons'' \cite{hoc09}.  At the same time, rejection rates are high in most journals
\cite{aar08,roh09}, and when not invited to submit a revision, authors must start the whole
process all over again.  Initiatives to reduce time from submission to
publication have emerged across the scientific community. Rohr et al.
\cite{roh09} called for the recycling and reuse of peer-reviews:
by attaching previous reviews and detailed replies to a new submission, both
the editor and the referees can gauge the work done on the manuscript, and
perhaps evaluate it with less prejudice. In a similar way, the \emph{Peerage
of Science} initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can
achieve the same goal of reducing the time spent in review. By putting a
manuscript for open comments and criticisms, the authors will receive valuable
feedback and can improve the version which will be submitted. With a rich
enough community of scientists depositing preprints, and commenting on them,
the process of an open pre- review can become widespread and will overall
increase the quality of first submissions.

\begin{figure}[ht!] \centering\includegraphics[width=0.50\textwidth]{map.pdf}
\caption { It can take several months, and even a few years, before a submitted
paper is officially published and citable.  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map} \end{figure}
% Dom : Do you have a reference for the stats in the legend?

Public preprint servers offer a much fairer way to establish intellectual
priority by making the work available when done. Some manuscripts will spend
much more time in the review process than others. Surprisingly, there is a
perception in biology that public preprints make it easier to steal ideas, as if
a scientific idea was only complete once published in a peer-reviewed journal
\cite{gin11}. Ironically, mathematicians and physicists have embraced arXiv in
part to establish priority in a fair way \cite{gin11,cal12}.
% Dom: That's a very interesting statement, if not the best one to justify preprints. Could you develop just a bit more the argument?

Prepublication reviews by a small network of colleagues is an important part of
the scientific process, attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the scientific process can be open
as soon as possible.
% Dom: I do accept this argument, although I have doubts it makes any difference. Just think about the number of comments posted on PLoS One...

@@

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, Figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually
be reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee % What about corrections? There are exchanges between referees and the authors?
response. F1000Research works closely with data providers to integrate raw
data to the paper. For instance, upon submitting a paper, authors are asked to
upload their data, which are then integrated in \emph{e.g.} FigShare widgets,
the DOI of which are given in the paper when the data are first mentionned.
The licensing of the data is similar to the one used by FigShare, meaning that
the articles are free to access, and can be redistributed readily. By putting
much effort in integrating data to the paper, F1000Research is working to make
science more reproducible and open.

\subsection{GitHub}

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous contributers to work asynchronously on the same project, often in parallel branches, all of which can be effortlessly merged and version controlled. Although created primarily for software development, git is ideal for academic research since it provides a way to version control and collaborate on every step of the manuscript development process, from data manipulation and analysis to writing and revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then merge their changes into a master branch.This takes the preprint process to an entirely new level, where the entire writing process is open from the beginning.
% I would avoid using the term forking since it is not a central piece of the section. It is one specific way of collaborating on a github repo when everyone does not have rights to it. People could just as easily work on a single branch, all with write access, either directly to the master branch or to individual branches that could be merged. So in other words, we should keep the language general so people don't get bogged down with trying to understand what a fork is.
% also I changed one author clones, then submits pull request langauge a bit since the main point of using git is that we can all work simultaneously unlike svn.
% I took out "These techniques are widely used in open source software development." because git can also be used for closed source software, offering the same advantages, where the repositories are private. Paid git accounts allow for private repos.

@@

% I added the following three paragraphs to the conclusion, offering also a discussion on the on going explosion of different publication models. Take what you want and remove what you don't need.
% I also had the feeling we might want to include a single paragraph, very short, calling that evaluations (academia, grant agencies etc...) looking at publication records or report of scientific activities start to view preprints as a significant and sound scientific contribution. One of the most important benefit is again the speed of the publication process. Most grants are for three years or less. It makes it hard to have a published paper within the limits of the grant. Preprints could however speed the process and show that contributions were made during the time of the grant. 
The ongoing discussions on the publication process, peer-reviewing and alternative publication models are all symptoms of the current uneasiness of scientists with the ever growing obsession with bibliographic metrics such as the impact factor \cite{Fisher2012}. There is pressure on researchers, wether they are junior looking for an academic position or senior trying to secure funding for their team, to orient their publication strategy in order to maximize their number of publications and total citations. A well-known consequence is to submit manuscripts first to the most prestigious journals, and then resubmit to lower level journals as they are rejected. The numerous negative impacts of such behavior have been discussed in depth \cite{hoc09}. They all contribute somehow in a general slowing down of scientific progress. Research activities and the publication process are drifting away from their fundamental object, namely the diffusion of novel scientific discoveries.

Developing a preprint culture in biology is not a final solution to the impact factor inflation, it might however reduce significantly its negative consequences. Preprints speed up publication, openly, with no judgement on pertinence and originality. Paradoxically, because peer-reviewing is conducted following this step in the process of a scientific communication, it might bring it closer to its fundamental objective. The role of peer-reviewing is to judge the scientific quality of a study, the match between its hypotheses, methods, results and conclusions. The peer review process is the first barrier against fraudulent or bad quality science, susceptible to impede scientific progress. It is however now dominated by subjective evaluations, where trendy contributions are promoted to increase the immediacy impact of the journal. Most journals have in their instructions for reviewers a section devoted to the evaluation of suitability for the journal, a task usually incumbent to editors. Preprints are by definition not evaluated, neither for their quality nor their pertinence. Technically, the difference between a preprint and a traditional publication should be that the latter as the approval stamp by pairs on its quality. The relevance of a study, its contribution to science, should only be judged post-publication by many more readers than the typical two-four anonymous reviewers. With a such a shift in the diffusion strategy, the role of traditional journals and their editors would be to showcase scientific discoveries for specialized readership. Such a process should improve even further their diffusion, not impede it.

The current opening of the publication process we are currently witnessing with the arrival of new preprint systems and innovative publication models such as F1000, PLoS One, Nature Scientific Reports, might significantly bypass the inflation of bibliographic metrics. They not only promote scientific community by changing the focus of the peer-reviewing, they also favor the publication of results that might otherwise be rejected because of low interest, negative results, descriptive and quantitative studies and datasets. One of drawback of making publication easier is obviously the proliferation of studies of uneven quality. A tradeoff between the intensity of the peer-review filtering and the benefits to science has been hypothesized \cite{Aarssen2012}. With increasingly stringent peer reviewing, the quality of published papers and easiness of finding discoveries might increase, at the cost of censorship, increasing load on authors and reviewers and time for publication. Preprints are simply bypassing this model, for what we believe is the progress of science: they speed up the dissemination of scientific discoveries, impede censorship and put on reader's shoulders the responsibility to judge originality and pertinence.

% Fisher, J, Ritchie, E.G., Hanspach, J. (2012). Academia's obsession with quantity. Trends in Ecology and Evolution, 27, 473-474.
% Aarssen, L. (2012). Are peer-review filters optimal for the progress of science in ecology and evolution? Ideas in Ecology and Evolution, 5, 9–12.
